---
title: "Nonlinear Mixed-Effects Model Development and Simulation Using nlmixr and Related R Open-Source Packages"
author:
  - Matthew Fidler, Novartis
  - Justin J. Wilkins, Occams
  - Richard Hooijmaijers, LAP\&P
  - Teun M. Post, LAP&P
  - Rik Schoemaker, Occams
  - Mirjam Trame, Novartis
  - Yuan Xiong, Certara
  - Wenping Wang, Novartis
  date: "`r Sys.Date()`"

output:
  md_document:
    variant: markdown_github
    toc: true
    toc_depth: 3
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  out.width = "100%",
  fig.path = "figure/README-"
  )
```

This vignette includes an updated and simplified version of the code used for the tutorial published in_Clinical Pharmacology & Therapeutics: Pharmacometrics and Systems Pharmacology_ in 2019 [https://doi.org/10.1002/psp4.12445](https://doi.org/10.1002/psp4.12445). It includes a number of updates to keep the examples working with current versions of `nlmixr2` and `rxode2`. Workflow is, on the whole, not different, although some commands and parameters have changed. 

```{r libraries}

library(gridExtra)
library(rxode2)
library(MASS)
library(data.table)
library(ggplot2)
library(nlmixr2)
library(xpose)
library(xpose.nlmixr2)
library(tibble)

```

## Setting up the data 

The first step is to simulate the data we will need to run the examples. For this tutorial, we'll use a two-compartmental PK model with some correlation between parameters, and covariate effects of body weight and sex. 

```{r simulate_data, cache=T}
mod <- rxode2({
  k10 = CL/V2
  k12 = Q/V2
  k21 = Q/V3
  d/dt(depot) =-KA*depot;
  d/dt(centr) = KA*depot - k10*centr - k12*centr + k21*peri;
  d/dt(peri)  =                        k12*centr - k21*peri;
  C2 = centr/V2;
  C3 = peri/V3;
  cp = C2 
})

theta <- c(TKA=1.05, TCL=0.121, TV2=1.939,
           TQ=0.282, TV3=5.65)

omegaCor <- matrix(c(1,    0.5, 0.25, 0.1,  0,
                     0.5,  1,   0.5,  0.1,  0,
                     0.25, 0.5, 1,    0.1,  0,
                     0.1,  0.1, 0.1,  1,    0,
                     0,    0,   0,    0,    1), 
                   dimnames=list(NULL,c("eta.CL",
                                        "eta.V2",
                                        "eta.V3", 
                                        "eta.Q", 
                                        "eta.KA")), nrow=5)

iiv.sd <- c(0.25, 0.25, 0.25, 0.3, 0.3) ## SDs of model parameters

iiv <- iiv.sd %*% t(iiv.sd)
omega <- iiv * omegaCor  # covariance matrix

sigma <- diag(1)*0.1
dimnames(sigma) <- list(NULL, c("cp.err"))

set.seed(740727)

mv <- mvrnorm(40, rep(0, dim(omega)[1]), omega) # Sample from covariance matrix

# Combine population parameters with IIV
params.all <-
  data.table(
    "ID" = seq(1:40),
    "CL" = theta['TCL'] * exp(mv[, 1]),
    "V2" = theta['TV2'] * exp(mv[, 2]),
    "V3" = theta['TV3'] * exp(mv[, 3]),
    "Q"  = theta['TQ']  * exp(mv[, 4]),
    "KA" = theta['TKA'] * exp(mv[, 5]),
    "WT" = round(rnorm(40,70,15)),
    "SEX" = rbinom(n = 40, prob = 0.5, size =1)
  )
#set the doses (looping through the 4 doses)
params.all[, AMT := 1200]

params.all$CL <- params.all$CL * (params.all$WT/70)^0.75
params.all$V2 <- params.all$V2 * (1 - 0.2 * params.all$SEX)

s = lapply(1:40, function(i) {
  # selects the parameters associated with the subject to be simulated
  params <- params.all[i]
  # creates an eventTable with 7 doses every 24 hours
  ev <- eventTable()
  ev$add.dosing(
    dose = params$AMT,
    nbr.doses = 28,
    dosing.to = 1,
    dosing.interval = 24,
    rate = NULL,
    start.time = 0
  )

  smp <- c(round(runif(1, 0, 1),3),
           round(runif(1, 1, 3),3),
           round(runif(1, 3, 6), 3),
           round(runif(1, 6, 12), 3),
           round(runif(1, 18, 23.9), 3),
           round(runif(1, 168, 169),3),
           round(runif(1, 169, 171),3),
           round(runif(1, 171, 180),3),
           round(runif(1, 188, 191.9),3))

  ev$add.sampling(smp)

  x <- as.data.table(mod$run(params, ev, seed=740727))
  x$rv <- rnorm(nrow(x), 0, 0.075)
  x$DV <- round(x$cp * (1 + x$rv),1)
  x$ID <- i

  x[, names(params) := params]
})

sim <- as.data.table(do.call("rbind", s))

setnames(sim, "time", "TIME")

Dose <- expand.grid(TIME = seq(0, 7 * 24, 24), ID = params.all$ID, DV=0)
Dose <- data.table(merge(Dose, params.all, by = "ID"))
Dose[, EVID := 101]

sim[, EVID := 0]
sim[, AMT := 0]

sim <- sim[,c("ID","TIME","DV","WT","SEX","AMT","EVID")]
dat <- rbind(sim, Dose[,c("ID","TIME","DV","WT","SEX","AMT","EVID")])
setkey(dat, ID, TIME)
```

Here's what the data look like, stratified by day.

```{r plot, cache=T, fig.width=8, fig.height=5}
dat$Day <- "Day 1"
dat$Day[dat$TIME>167] <- "Day 8"

ggplot(subset(data.frame(dat), EVID==0), aes(TIME, DV)) + geom_point(col="#1F4E79") +
  geom_line(aes(group=ID), col="#1F4E79") +
  scale_x_continuous("Time (h)") +
  scale_y_log10("Concentration (mg/L)") +
  facet_wrap(~ Day, scales="free_x")
```


## Model exploration

### One-compartmental, using NLME

We start our model exploration using a one-compartmental model, fitted using the NLME method. This is known to be terrible, but let's see just how terrible. We will request NPDE and CWRES by specifying `table=tableControl(cwres=TRUE, npde=TRUE)`. 

```{r 1-cpt-nlme, cache=T}
model.1cpt.cf <- function() {
  ini({
    tka <- log(1.5)
    tcl <- log(1.5)
    tv  <- log(3)
    eta.ka ~ 0.1
    eta.cl ~ 0.1
    eta.v ~ 0.1
    add.err <- 0.01
  })
  model({
    ka <- exp(tka + eta.ka)
    cl <- exp(tcl + eta.cl)
    v <- exp(tv + eta.v)
    linCmt() ~ add(add.err)
  })
}

fit.1cpt.cf.nlme <- nlmixr2(model.1cpt.cf, dat, est="nlme", table=tableControl(cwres=TRUE, npde=TRUE))
```

We can see a snapshot of the results by using `print`.

```{r 1-cpt-nlme-results, cache=T}
print(fit.1cpt.cf.nlme)
```

Now let's fit the same model using SAEM, and ordinary differential equations (ODEs).

```{r 1cpt-saem, cache=T}
model.1cpt.ode <- function() {
  ini({
    tka <- log(1.5)
    tcl <- log(1.5)
    tv  <- log(3)
    eta.ka ~ 1
    eta.cl ~ 1
    eta.v ~ 1
    prop.err <- 0.1
  })
  model({
    ka <- exp(tka + eta.ka)
    cl <- exp(tcl + eta.cl)
    v <- exp(tv + eta.v)
    d/dt(depot) = -ka * depot
    d/dt(center) = ka * depot - cl / v * center
    cp = center / v
    cp ~ prop(prop.err)
  })
}

fit.1cpt.ode.saem <- nlmixr2(model.1cpt.ode, dat, est="saem", table=tableControl(cwres=TRUE, npde=TRUE), seed=740727)
```

The default for SAEM is to run 500 iterations. 

```{r 1cpt-saem-results, cache=T}
print(fit.1cpt.ode.saem)
```

From the output, we see the objective function - calculated using FOCEI - is **`r round(as.numeric(fit.1cpt.ode.saem$objDf[1]),3)`**. Let's look at some diagnostic plots.

```{r 1cpt-diagnostics, cache=T, fig.width=8, fig.height=8}
xp.1cpt.ode.saem <- xpose_data_nlmixr2(fit.1cpt.ode.saem, xp_theme = theme_xp_nlmixr2())

xp1 <- dv_vs_pred(xp.1cpt.ode.saem, title = "DV vs PRED",
                  subtitle = NULL, caption = NULL) +
  coord_cartesian(ylim=c(0,1000), xlim=c(0,1000)) +
  scale_x_continuous("Population predictions") +
  scale_y_continuous("Observed concentrations")

xp2 <- dv_vs_ipred(xp.1cpt.ode.saem, title = "DV vs IPRED",
                   subtitle = NULL, caption = NULL) +
  coord_cartesian(ylim=c(0,1000), xlim=c(0,1000))+
  scale_x_continuous("Individual predictions") +
  scale_y_continuous("Observed concentrations")

xp3 <- res_vs_idv(xp.1cpt.ode.saem, res = "CWRES", title = "CWRES vs time",
                  subtitle = NULL, caption = NULL) +
  coord_cartesian(ylim=c(-3.5,3.5)) +
  scale_x_continuous("Time (h)") +
  scale_y_continuous("Conditional weighted residuals")

xp4 <- res_vs_pred(xp.1cpt.ode.saem, res = "CWRES", title = "CWRES vs PRED",
                   subtitle = NULL, caption = NULL)+
  coord_cartesian(ylim=c(-3.5,3.5))+
  scale_x_continuous("Population predictions") +
  scale_y_continuous("Conditional weighted residuals")

grid.arrange(xp1, xp2, xp3, xp4, nrow=2)
```

That's pretty horrible. What about a VPC? `nlmixr` makes VPCs quick and easy.

```{r 1cpt-vpc, cache=T, fig.width=8, fig.height=5}
bins <- c(0, 3, 12, 24, 168, 171, 180, 192)
vpc.1cpt <- nlmixr2::vpcPlot(fit.1cpt.ode.saem, n=400,show=list(obs_dv=T),
                        bins=bins);
vpc.1cpt +
    scale_x_continuous("Time (h)") +
    scale_y_continuous("Concentration")
```

Still horrible. 

We know that the true model is two-compartmental, because that's how we simulated it. Let's try that. 

```{r 2cpt-saem, cache=T}
model.2cpt.ode <- function() {
  ini({
    tka <- log(1.05)
    tcl <- log(0.121)
    tv2  <- log(1.939)
    tv3  <- log(5.65)
    tq   <- log(0.282)
    eta.ka ~ 0.1
    eta.cl ~ 0.1
    eta.v2 ~ 0.1
    eta.v3 ~ 0.1
    eta.q ~ 0.1
    add.err <- 75
  })
  model({
    ka <- exp(tka + eta.ka)
    cl <- exp(tcl + eta.cl)
    v2 <- exp(tv2 + eta.v2)
    v3 <- exp(tv3 + eta.v3)
    q  <- exp(tq + eta.q)
    d/dt(depot) = -ka * depot
    d/dt(center) = ka * depot - cl / v2 * center + q/v3 * periph - q/v2 * center
    d/dt(periph) = q/v2 * center - q/v3 * periph
    cp = center / v2
    cp ~ add(add.err)
  })
}

fit.2cpt.ode.saem <- nlmixr2(model.2cpt.ode, dat, est="saem", table=tableControl(cwres=TRUE, npde=TRUE), seed=740727)
```

```{r 2cpt-saem-results, cache=T}
print(fit.2cpt.ode.saem)
```

We have reduced the OFV from **`r round(as.numeric(fit.1cpt.ode.saem$objDf[1]),3)`** to **`r round(as.numeric(fit.2cpt.ode.saem$objDf[1]),3)`**, a drop of **`r round(as.numeric(fit.1cpt.ode.saem$objDf[1])-as.numeric(fit.2cpt.ode.saem$objDf[1]),3)`**! Progress. 

What about the residual error? We're using an additive model. What happens when we try proportional?

```{r 2cpt-saem-prop, cache=T}
model.2cptp.ode <- function() {
  ini({
    tka <- log(1.05)
    tcl <- log(0.121)
    tv2  <- log(1.939)
    tv3  <- log(5.65)
    tq   <- log(0.282)
    eta.ka ~ 0.1
    eta.cl ~ 0.1
    eta.v2 ~ 0.1
    eta.v3 ~ 0.1
    eta.q ~ 0.1
    prop.err <- 0.075
  })
  model({
    ka <- exp(tka + eta.ka)
    cl <- exp(tcl + eta.cl)
    v2 <- exp(tv2 + eta.v2)
    v3 <- exp(tv3 + eta.v3)
    q  <- exp(tq + eta.q)
    d/dt(depot) = -ka * depot
    d/dt(center) = ka * depot - cl / v2 * center + q/v3 * periph - q/v2 * center
    d/dt(periph) = q/v2 * center - q/v3 * periph
    cp = center / v2
    cp ~ prop(prop.err)
  })
}

fit.2cptp.ode.saem <- nlmixr2(model.2cptp.ode, dat, est="saem", table=tableControl(cwres=TRUE, npde=TRUE), seed=740727)
```

```{r 2cpt-saem-prop-results, cache=T}
print(fit.2cptp.ode.saem)
```

A further drop of **`r round(as.numeric(fit.2cpt.ode.saem$objDf[1])-as.numeric(fit.2cptp.ode.saem$objDf[1]),3)`** points!

```{r 2cpt-diagnostics, cache=T, fig.width=8, fig.height=8}
xp.2cpt.ode.saem <- xpose_data_nlmixr2(fit.2cptp.ode.saem, xp_theme = theme_xp_nlmixr2())

xp1 <- dv_vs_pred(xp.2cpt.ode.saem, title = "DV vs PRED",
                  subtitle = NULL, caption = NULL) +
  coord_cartesian(ylim=c(0,1000), xlim=c(0,1000)) +
  scale_x_continuous("Population predictions") +
  scale_y_continuous("Observed concentrations")

xp2 <- dv_vs_ipred(xp.2cpt.ode.saem, title = "DV vs IPRED",
                   subtitle = NULL, caption = NULL) +
  coord_cartesian(ylim=c(0,1000), xlim=c(0,1000))+
  scale_x_continuous("Individual predictions") +
  scale_y_continuous("Observed concentrations")

xp3 <- res_vs_idv(xp.2cpt.ode.saem, res = "CWRES", title = "CWRES vs time",
                  subtitle = NULL, caption = NULL) +
  coord_cartesian(ylim=c(-3.5,3.5)) +
  scale_x_continuous("Time (h)") +
  scale_y_continuous("Conditional weighted residuals")

xp4 <- res_vs_pred(xp.2cpt.ode.saem, res = "CWRES", title = "CWRES vs PRED",
                   subtitle = NULL, caption = NULL)+
  coord_cartesian(ylim=c(-3.5,3.5))+
  scale_x_continuous("Population predictions") +
  scale_y_continuous("Conditional weighted residuals")

grid.arrange(xp1, xp2, xp3, xp4, nrow=2)
```

Better.

```{r 2cpt-vpc, cache=T, fig.width=8, fig.height=5}
bins <- c(0, 3, 12, 24, 168, 171, 180, 192)
vpc.2cpt <- nlmixr2::vpcPlot(fit.2cptp.ode.saem, n=400,show=list(obs_dv=T),
                        bins=bins);
vpc.2cpt +
    scale_x_continuous("Time (h)") +
    scale_y_continuous("Concentration")
```

Much better!

<!-- For good measure, what does a 3-compartmental model look like? -->

<!-- ```{r 3cpt-saem, cache=T} -->
<!-- model.3cpt.ode <- function() { -->
<!--   ini({ -->
<!--     tka <- log(1.42) -->
<!--     tcl <- log(0.044) -->
<!--     tv2  <- log(2) -->
<!--     tv3  <- log(10) -->
<!--     tv4  <- log(10) -->
<!--     tq2   <- log(0.5) -->
<!--     tq3   <- log(0.5) -->
<!--     eta.ka ~ 0.1 -->
<!--     eta.cl ~ 0.1 -->
<!--     eta.v2 ~ 0.1 -->
<!--     eta.v3 ~ 0.1 -->
<!--     eta.v4 ~ 0.1 -->
<!--     eta.q2 ~ 0.1 -->
<!--     eta.q3 ~ 0.1 -->
<!--     prop.err <- 0.075 -->
<!--   }) -->
<!--   model({ -->
<!--     ka <- exp(tka + eta.ka) -->
<!--     cl <- exp(tcl + eta.cl) -->
<!--     v2 <- exp(tv2 + eta.v2) -->
<!--     v3 <- exp(tv3 + eta.v3) -->
<!--     v4 <- exp(tv4 + eta.v4) -->
<!--     q2  <- exp(tq2 + eta.q2) -->
<!--     q3  <- exp(tq3 + eta.q3) -->
<!--     d/dt(depot) = -ka * depot -->
<!--     d/dt(center) = ka * depot - cl / v2 * center + q2/v3 * periph1 - q2/v2 * center + q3/v4 * periph2 - q3/v2 * center -->
<!--     d/dt(periph1) = q2/v2 * center - q2/v3 * periph1 -->
<!--     d/dt(periph2) = q3/v2 * center - q3/v4 * periph2 -->
<!--     cp = center / v2 -->
<!--     cp ~ prop(prop.err) -->
<!--   }) -->
<!-- } -->

<!-- fit.3cpt.ode.saem <- nlmixr2(model.3cpt.ode, dat, est="saem", table=tableControl(cwres=TRUE, npde=TRUE), seed=740727) -->
<!-- ``` -->

<!-- ```{r 3cpt-saem-prop-results, cache=T} -->
<!-- print(fit.3cpt.ode.saem) -->
<!-- ``` -->

<!-- Looks better, doesn't it? Although we know it's not true. Let's tabulate the results so far. -->

<!-- ```{r modcmp1, cache=T} -->
<!-- otab <- data.frame(Model = c(1, 2, 3, 4), -->
<!--                    RelativeTo = c("-", 1, 2, 3), -->
<!--                    Description = c("1-cpt, additive residual error", -->
<!--                              "2-cpt, additive residual error", -->
<!--                              "2-cpt, proportional residual error", -->
<!--                              "3-cpt, proportional residual error"), -->
<!--                    OFV = c(fit.1cpt.ode.saem$value, fit.2cpt.ode.saem$value, fit.2cptp.ode.saem$value, fit.3cpt.ode.saem$value), -->
<!--                    dOFV = round(c(0, fit.2cpt.ode.saem$value-fit.1cpt.ode.saem$value, -->
<!--                             fit.2cptp.ode.saem$value-fit.2cpt.ode.saem$value, fit.3cpt.ode.saem$value-fit.2cptp.ode.saem$value),3), -->
<!--                    AIC = c(AIC(fit.1cpt.ode.saem), AIC(fit.2cpt.ode.saem), AIC(fit.2cptp.ode.saem), AIC(fit.3cpt.ode.saem)), -->
<!--                    dAIC = round(c(0, AIC(fit.2cpt.ode.saem)-AIC(fit.1cpt.ode.saem), -->
<!--                             AIC(fit.2cptp.ode.saem)-AIC(fit.2cpt.ode.saem), AIC(fit.3cpt.ode.saem)-AIC(fit.2cptp.ode.saem)),3), -->
<!--                    BIC = c(BIC(fit.1cpt.ode.saem), BIC(fit.2cpt.ode.saem), BIC(fit.2cptp.ode.saem), BIC(fit.3cpt.ode.saem)), -->
<!--                    dBIC = round(c(0, BIC(fit.2cpt.ode.saem)-BIC(fit.1cpt.ode.saem), -->
<!--                             BIC(fit.2cptp.ode.saem)-BIC(fit.2cpt.ode.saem), BIC(fit.3cpt.ode.saem)-BIC(fit.2cptp.ode.saem)),3)) -->

<!-- knitr::kable( -->
<!--   otab, longtable = TRUE, booktabs = TRUE) -->
<!-- ``` -->

<!-- For every metric except BIC, the 3-compartment model looks better, although we know it is not the correct model. Let's run with the 2-compartment model for now.  -->

## Covariates

Now let's try some covariates! We know weight and sex are important, since we had them in the simulation model. Weight first...

```{r 2cptwt, cache=T}
dat$lnWT <- log(dat$WT/70)

model.2cpt.ode.wtcl <- function() {
  ini({
    tka <- log(1.05)
    tcl <- log(0.121)
    tv2  <- log(1.939)
    tv3  <- log(5.65)
    tq   <- log(0.282)
    wteff <- 0.75
    eta.ka ~ 0.2
    eta.cl ~ 0.2
    eta.v2 ~ 0.2
    eta.v3 ~ 0.2
    eta.q ~ 0.2
    prop.err <- 0.075
  })
  model({
    ka = exp(tka + eta.ka)
    cl = exp(tcl + wteff*lnWT + eta.cl)
    v2 = exp(tv2 + eta.v2)
    v3 = exp(tv3 + eta.v3)
    q  = exp(tq + eta.q)
    d/dt(depot) = -ka * depot
    d/dt(center) = ka * depot - cl / v2 * center + q/v3 * periph - q/v2 * center
    d/dt(periph) = q/v2 * center - q/v3 * periph
    cp = center / v2
    cp ~ prop(prop.err)
  })
}

fit.2cpt.ode.wtcl.saem <- nlmixr2(model.2cpt.ode.wtcl, dat, est="saem", table=tableControl(cwres=TRUE, npde=TRUE))
```

```{r 2cptsex, cache=T}
model.2cpt.ode.sexv2 <- function() {
  ini({
    tka <- log(1.15)
    tcl <- log(0.12)
    tv2  <- log(1.82)
    tv3  <- log(5.73)
    tq   <- log(0.278)
    sexeff <- -0.01
    eta.ka ~ 0.06626505
    eta.cl ~ 0.09320262
    eta.v2 ~ 0.05237126
    eta.v3 ~ 0.008968923
    eta.q ~ 0.08214272
    prop.err <- 0.075
  })
  model({
    ka = exp(tka + eta.ka)
    cl = exp(tcl + eta.cl)
    v2 = exp(tv2 + sexeff*(SEX) + eta.v2)
    v3 = exp(tv3 + eta.v3)
    q  = exp(tq + eta.q)
    d/dt(depot) = -ka * depot
    d/dt(center) = ka * depot - cl / v2 * center + q/v3 * periph - q/v2 * center
    d/dt(periph) = q/v2 * center - q/v3 * periph
    cp = center / v2
    cp ~ prop(prop.err)
  })
}

fit.2cpt.ode.sexv2.saem <- nlmixr2(model.2cpt.ode.sexv2, dat, est="saem", table=tableControl(cwres=TRUE, npde=TRUE))
```

```{r 2cptwtsex, cache=T}
model.2cpt.ode.wtcl.sexv2 <- function() {
  ini({
    tka <- log(1.05)
    tcl <- log(0.121)
    tv2  <- log(1.939)
    tv3  <- log(5.65)
    tq   <- log(0.282)
    wteff  <- 0.75
    sexeff <- -0.2
    eta.ka ~ 0.06626505
    eta.cl ~ 0.09320262
    eta.v2 ~ 0.05237126
    eta.v3 ~ 0.008968923
    eta.q ~ 0.08214272
    prop.err <- 0.075
  })
  model({
    ka <- exp(tka + eta.ka)
    cl <- exp(tcl + wteff*lnWT + eta.cl)
    v2 <- exp(tv2 + sexeff*(SEX) + eta.v2)
    v3 <- exp(tv3 + eta.v3)
    q  <- exp(tq + eta.q)
    d/dt(depot) = -ka * depot
    d/dt(center) = ka * depot - cl / v2 * center + q/v3 * periph - q/v2 * center
    d/dt(periph) = q/v2 * center - q/v3 * periph
    cp = center / v2
    cp ~ prop(prop.err)
  })
}

## broke.
fit.2cpt.ode.wtcl.sexv2.saem <- nlmixr2(model.2cpt.ode.wtcl.sexv2, dat, est="saem", table=tableControl(cwres=TRUE, npde=TRUE))
```

Time to tabulate again...

```{r covtab, cache=T}
otab <- data.frame(Model = c(1, 2, 3, 4, 5, 6, 7),
                   RelativeTo = c("-", 1, 2, 3, 3, 3, 5),
                   Description = c("1-cpt, add residual error",
                             "2-cpt, add residual error",
                             "2-cpt, prop residual error (base)",
                             "3-cpt, prop residual error",
                             "Base with WT-CL",
                             "Base with SEX-V2",
                             "Base with WT-CL, SEX-V2 (final)"),
                   OFV = c(fit.1cpt.ode.saem$value, fit.2cpt.ode.saem$value, fit.2cptp.ode.saem$value, fit.3cpt.ode.saem$value,
                           fit.2cpt.ode.wtcl.saem$value, fit.2cpt.ode.sexv2.saem$value, fit.2cpt.ode.wtcl.sexv2.saem$value),
                   dOFV = round(c(0,
                                  fit.2cpt.ode.saem$value-fit.1cpt.ode.saem$value,
                                  fit.2cptp.ode.saem$value-fit.2cpt.ode.saem$value,
                                  fit.3cpt.ode.saem$value-fit.2cptp.ode.saem$value,
                                  fit.2cpt.ode.wtcl.saem$value-fit.2cptp.ode.saem$value,
                                  fit.2cpt.ode.sexv2.saem$value-fit.2cptp.ode.saem$value,
                                  fit.2cpt.ode.wtcl.sexv2.saem$value-fit.2cpt.ode.wtcl.saem$value),3),
                   AIC = c(AIC(fit.1cpt.ode.saem), AIC(fit.2cpt.ode.saem), AIC(fit.2cptp.ode.saem), AIC(fit.3cpt.ode.saem),
                           AIC(fit.2cpt.ode.wtcl.saem), AIC(fit.2cpt.ode.sexv2.saem), AIC(fit.2cpt.ode.wtcl.sexv2.saem)),
                   dAIC = round(c(0,
                                  AIC(fit.2cpt.ode.saem)-AIC(fit.1cpt.ode.saem),
                                  AIC(fit.2cptp.ode.saem)-AIC(fit.2cpt.ode.saem),
                                  AIC(fit.3cpt.ode.saem)-AIC(fit.2cptp.ode.saem),
                                  AIC(fit.2cpt.ode.wtcl.saem)-AIC(fit.2cptp.ode.saem),
                                  AIC(fit.2cpt.ode.sexv2.saem)-AIC(fit.2cptp.ode.saem),
                                  AIC(fit.2cpt.ode.wtcl.sexv2.saem)-AIC(fit.2cpt.ode.wtcl.saem)),3),
                   BIC = c(BIC(fit.1cpt.ode.saem), BIC(fit.2cpt.ode.saem), BIC(fit.2cptp.ode.saem), BIC(fit.3cpt.ode.saem),
                           BIC(fit.2cpt.ode.wtcl.saem), BIC(fit.2cpt.ode.sexv2.saem), BIC(fit.2cpt.ode.wtcl.sexv2.saem)),
                   dBIC = round(c(0,
                                  BIC(fit.2cpt.ode.saem)-BIC(fit.1cpt.ode.saem),
                                  BIC(fit.2cptp.ode.saem)-BIC(fit.2cpt.ode.saem),
                                  BIC(fit.3cpt.ode.saem)-BIC(fit.2cptp.ode.saem),
                                  BIC(fit.2cpt.ode.wtcl.saem)-BIC(fit.2cptp.ode.saem),
                                  BIC(fit.2cpt.ode.sexv2.saem)-BIC(fit.2cptp.ode.saem),
                                  BIC(fit.2cpt.ode.wtcl.sexv2.saem)-BIC(fit.2cpt.ode.wtcl.saem)),3))

knitr::kable(
  otab, longtable = TRUE, booktabs = TRUE)
```

As we can see, the model with both weight and sex fits the best...

```{r 2cpt-cov-diagnostics, cache=T, fig.width=8, fig.height=8}
xp.2cpt.cov.ode.saem <- xpose_data_nlmixr2(fit.2cpt.ode.wtcl.sexv2.saem, xp_theme = theme_xp_nlmixr2())

xp1 <- dv_vs_pred(xp.2cpt.cov.ode.saem, title = "DV vs PRED",
                  subtitle = NULL, caption = NULL) +
  coord_cartesian(ylim=c(0,1000), xlim=c(0,1000)) +
  scale_x_continuous("Population predictions") +
  scale_y_continuous("Observed concentrations")

xp2 <- dv_vs_ipred(xp.2cpt.cov.ode.saem, title = "DV vs IPRED",
                   subtitle = NULL, caption = NULL) +
  coord_cartesian(ylim=c(0,1000), xlim=c(0,1000))+
  scale_x_continuous("Individual predictions") +
  scale_y_continuous("Observed concentrations")

xp3 <- res_vs_idv(xp.2cpt.cov.ode.saem, res = "CWRES", title = "CWRES vs time",
                  subtitle = NULL, caption = NULL) +
  coord_cartesian(ylim=c(-3.5,3.5)) +
  scale_x_continuous("Time (h)") +
  scale_y_continuous("Conditional weighted residuals")

xp4 <- res_vs_pred(xp.2cpt.cov.ode.saem, res = "CWRES", title = "CWRES vs PRED",
                   subtitle = NULL, caption = NULL)+
  coord_cartesian(ylim=c(-3.5,3.5))+
  scale_x_continuous("Population predictions") +
  scale_y_continuous("Conditional weighted residuals")

grid.arrange(xp1, xp2, xp3, xp4, nrow=2)
```

```{r 2cpt-cov-vpc, cache=T, fig.width=8, fig.height=5}
bins <- c(0, 3, 12, 24, 168, 171, 180, 192)
vpc.2cpt.cov <- nlmixr2::vpcPlot(fit.2cpt.ode.wtcl.sexv2.saem, n=400,show=list(obs_dv=T),
                        bins=bins);
vpc.2cpt.cov +
    scale_x_continuous("Time (h)") +
    scale_y_continuous("Concentration")
```

And we get the following set of parameter estimates -

```{r partab, cache=T}
knitr::kable(
  fit.2cpt.ode.wtcl.sexv2.saem$parFixed, longtable = TRUE, booktabs = TRUE)
```

Not every parameter is spot on, but we haven't done too badly in recovering what we used for simulation.

## Simulations

We can use our model as a basis for simulation, by feeding it back into `rxode2`!

```{r simdes, cache=T}
ev <- eventTable(amount.units="mg", time.units="hours") %>%
    add.dosing(dose=600, nbr.doses=10, dosing.interval=24) %>%
    add.dosing(dose=2000, nbr.doses=5,
               start.time=240,dosing.interval=48) %>%
    add.sampling(0:480);
```

```{r sim1, cache=T}
sim.2cpt <- rxSolve(fit.2cpt.ode.saem, events=ev, nSub=500)

length(unique(sim.2cpt$sim.id))
head(sim.2cpt)

ssim <- as_tibble(sim.2cpt) %>%
  group_by(time) %>%
  dplyr::summarise(
    med = quantile(cp, 0.5),
    q025 = quantile(cp, 0.025),
    q05 = quantile(cp, 0.25),
    q25 = quantile(cp, 0.5),
    q75 = quantile(cp, 0.75),
    q95 = quantile(cp, 0.95),
    q975 = quantile(cp, 0.975)
  ) 
```

```{r sim1plot, cache=T, fig.height=5, fig.width=8}
ggplot(units::drop_units(ssim), aes(time, med)) +
  geom_ribbon(aes(time, ymax=q975, ymin=q025), fill="red", alpha=0.1) +
  geom_ribbon(aes(time, ymax=q95, ymin=q05), fill="red", alpha=0.1) +
  geom_ribbon(aes(time, ymax=q75, ymin=q25), fill="red", alpha=0.1) +
  geom_line(col="red") +
  scale_x_continuous("Time") +
  scale_y_continuous("Concentration")
  
```


